---
title: "US_Stock_Market"
author: "Anshul Singhal"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

###Loading the Packages which are required
```{r include = FALSE}
library(dplyr)
require(data.table)
library(haven)
library(zeallot)
library(lubridate)
library(plotly)
library(openxlsx)
library(readxl)
library(zoo)
library(scales)
library(geckor)
library(fredr)
library(tidyquant)
library(tidyr)
library(naniar)
library(tidyverse)
fredr_set_key("669c24e3e2c8136b6db6b3c978104993")
```

###Relevant Functions used in the Code
```{r}
"Multiple Line chart Function to draw line chart across time with the Macro economic variables"
plot_multiple_line_chart = function(dat1, title_type,y1_type,macro_flag = 0,type1 = "scatter") {
  if(macro_flag == 0){
    cols_taken = ncol(dat1)
  }else{
    cols_taken = ncol(dat1)-1
  }
  plot_y <- dat1 %>% plot_ly()
  for(i in c(2:ncol(dat1))) {
    if (i <= cols_taken) {
      x = plot_y %>% add_trace(x = ~DATE_START, y=dat1[[i]],  mode="markers",marker = list(size = 3), type = type1, name=colnames(dat1)[i], yaxis='y1')
    } else if(macro_flag != 0){
      x = plot_y %>% add_trace(x = ~DATE_START, y=dat1[[i]], mode="markers", type = "scatter", name=colnames(dat1)[i], yaxis='y2') 
    }
    plot_y = x
  }
  if(macro_flag != 0)
  {
      plot_y %>% 
      layout(title = paste0(title_type,"(LHS) vs Return (RHS)"),
         barmode   = 'relative', 
         xaxis     = list(title=''),
         margin    = list(l = 75, r = 75, b = 50, t = 50, pad = 4),
         xaxis     = list(title = ""),
         yaxis     = list(side = 'left', 
                       title = y1_type, 
                       showgrid = FALSE, 
                       zeroline = TRUE, 
                       color = 'steelblue'),
         yaxis2    = list(side = 'right', 
                       overlaying = "y", 
                       title = colnames(dat1)[ncol(dat1)], 
                       showgrid = TRUE, 
                       zeroline = FALSE,
                       ticksuffix = "%",
                       color = "#ffa500"),
         legend = list(traceorder = 'reversed',orientation = "h"))
  }
  else{
    plot_y %>% 
    layout(title = paste0(title_type," Data"),
       barmode   = 'relative', 
       xaxis     = list(title='Month'),
       margin    = list(l = 75, r = 75, b = 50, t = 50, pad = 4),
       xaxis     = list(title = ""),
       yaxis     = list(side = 'left', 
                       title = y1_type, 
                       showgrid = FALSE, 
                       zeroline = TRUE, 
                       color = 'steelblue'),
       legend = list(traceorder = 'reversed',orientation = "h"))
  }
}
"Plot Stacked Bar Chart over time"
plot_stacked_chart = function(dat1,col_name,grouped_column,title_name,y_axis_name){
  dat1["value"] = dat1[col_name]
  dat1["Type"] = dat1[grouped_column]
  stacked_plot = ggplot(dat1) +
        geom_bar(aes(x=DATE_START,y=value,fill=Type),
        stat='identity')+
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        labs(title = title_name, x = "Date", y = y_axis_name,color = "")
  ggplotly(stacked_plot)
}


"Descriptive Stats Function : count, mean,p1,p5, p25, p50, p75,p95,p99, std,skewness,kurtosis max, min"
stats = function(dat){
count_group = dat %>%
  summarise(across(everything(),funs(sum(!is.na(.))),.names = "count_{.col}"))
mean_group = dat %>%
  summarise(across(everything(), mean,na.rm = TRUE,.names = "mean_{.col}"))
p1_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.01),na.rm = TRUE,.names = "p01_{.col}"))
p5_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.05),na.rm = TRUE,.names = "p05_{.col}"))
p25_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.25),na.rm = TRUE,.names = "p25_{.col}"))
p50_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.5),na.rm = TRUE,.names = "p50_{.col}"))
p75_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.75),na.rm = TRUE,.names = "p75_{.col}"))
p95_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.95),na.rm = TRUE,.names = "p95_{.col}"))
p99_group = dat %>%
  summarise(across(everything(), quantile,probs = c(0.99),na.rm = TRUE,.names = "p95_{.col}"))
std_group = dat %>%
  summarise(across(everything(), sd,na.rm = TRUE,.names = "std_{.col}"))
skewness_group = dat %>%
  summarise(across(everything(), skewness, na.rm = TRUE,.names = "skew_{.col}"))
kurtosis_group = dat %>%
  summarise(across(everything(),kurtosis, na.rm = TRUE,.names = "kurt_{.col}"))
max_group = dat %>%
  replace(is.na(.), 0)%>%
  summarise(across(everything(),max,.names = "max_{.col}"))
min_group = dat %>%
  replace(is.na(.), 0)%>%
  summarise(across(everything(), min,.names = "min_{.col}"))
 return (list(count_group,mean_group,p1_group, p5_group, p25_group,p50_group, p75_group,p95_group,p99_group, std_group,skewness_group,kurtosis_group,max_group,min_group))
}

```

###Reading the Monthly Data and Name Mappings/Industry/Exchange (MSF New & Msenames)
```{r}
path = ""
mse_names = read.csv(paste0(path,"msenames.csv"))
monthly_data = fread(paste0(path,"msf_new2.csv"))
monthly_data = monthly_data%>%
  tibble()
```

##Getting industry and Exchange Mapping and Deflating to 2010 Dollar Terms
```{r}
mse_names_unique = mse_names%>%
  distinct(CUSIP,.keep_all= TRUE)

monthly_data_modified = monthly_data%>%
  merge(mse_names_unique[c("CUSIP","COMNAM","EXCHCD","SICCD")],by = "CUSIP",all.x = TRUE)%>%
  arrange(CUSIP,DATE)%>%
  drop_na(PRC)%>%
  filter(SHRCD %in% c(10,11))%>%
  mutate(DATE = as.Date(parse_date_time(DATE, orders = "Ymd")),MKTCAP = abs(PRC*SHROUT))

cpi_deflator <- 
   "CPIAUCNS" %>% 
  tq_get(get = "economic.data", from = "1925-01-01") %>% 
  select(DATE_START = date, INDEX  =  price)
cpi_deflator = cpi_deflator%>%
  mutate(MULT = cpi_deflator$INDEX[cpi_deflator$DATE_START == "2010-12-01"]/INDEX)

monthly_data_modified = monthly_data_modified%>%
  mutate(DATE_START = floor_date(DATE, unit = "month"))%>%
  merge(cpi_deflator[c("DATE_START","MULT")],by = "DATE_START",all.x = TRUE)%>%
mutate(MKTCAP_ADJ = MKTCAP*MULT,CATEGORY = if_else(EXCHCD %in% c(1,31),"NYSE",if_else(EXCHCD %in% c(2,32),"NYSE MKT",if_else(EXCHCD %in% c(3,33),"NASDAQ",if_else(EXCHCD %in% c(4,34),"ARCA","OTHER EXCHANGE")))))


```

##Count and Mean Plots for the Exchange wise Stocks over time(1925-2020)
```{r}
#Calculating the count and mean market cap
count_ticker = monthly_data_modified %>%
  group_by(CATEGORY,DATE_START)%>%
  summarise(across(c("CUSIP"),funs(sum(!is.na(.))),.names = "count_{.col}"))

mean_mcap = monthly_data_modified %>%
  group_by(CATEGORY,DATE_START)%>%
  summarise(across(c("MKTCAP_ADJ"), mean,na.rm = TRUE,.names = "mean_{.col}"))

#Stacked Bar chart plot
plot_stacked_chart(count_ticker,"count_CUSIP","CATEGORY","Total Stocks From 1925 Onwards By Exchange","Count")
plot_stacked_chart(mean_mcap,"mean_MKTCAP_ADJ","CATEGORY","Mean Overall Market Cap From 1925 Onwards By Exchange","Market Cap ($)")


```

##Count and Mean Plots for the Industry wise Stocks over time(1925-2020)
```{r}
##industry wise Plots and Analysis
#Mapping the industry
monthly_data_modified = monthly_data_modified%>%
  mutate(INDUSTRY = if_else(SICCD %in% c(1:999),"Agriculture, Forestry and Fishing",if_else(SICCD %in% c(1000:1499),"Mining",if_else(SICCD %in% c(1500:1799),"Construction",if_else(SICCD %in% c(2000:3999),"Manufacturing",if_else(SICCD %in% c(4000:4999),"Transportation and other Utilities",if_else(SICCD %in% c(5000:5199),"Wholesale Trade",if_else(SICCD %in% c(5200:5999),"Retail Trade",if_else(SICCD %in% c(6000:6799),"Finance, Insurance and Real Estate",if_else(SICCD %in% c(7000:8999),"Services",if_else(SICCD %in% c(9000:9999),"Public Administration","Others")))))))))))

#Calculating the count and mean market cap
count_ticker_industry = monthly_data_modified %>%
  group_by(INDUSTRY,DATE_START)%>%
  summarise(across(c("CUSIP"),funs(sum(!is.na(.))),.names = "count_{.col}"))

mean_mcap_industry = monthly_data_modified %>%
  group_by(INDUSTRY,DATE_START)%>%
  summarise(across(c("MKTCAP_ADJ"), mean,na.rm = TRUE,.names = "mean_{.col}"))

#Stacked Bar chart plot
plot_stacked_chart(count_ticker_industry,"count_CUSIP","INDUSTRY","Total Stocks From 1925 Onwards By Industry","Count")
plot_stacked_chart(mean_mcap_industry,"mean_MKTCAP_ADJ","INDUSTRY","Mean Overall Market Cap From 1925 Onwards By Industry","Market Cap ($)")


```

##Reading the FAMA FRENCH Risk Free Rate Data (RF)
```{r}
#Downloading and Reading the data
risk_free_data = read.csv("F-F_Research_Data_Factors.csv",skip = 2)
names(risk_free_data)[1] = "DATE_START"
risk_free_data = risk_free_data%>%
  mutate(DATE_START = as.Date(parse_date_time(DATE_START, orders = "Ym")))%>%
  mutate_if(is.character, as.numeric)

monthly_data_modified_risk_free = monthly_data_modified%>%
  merge(risk_free_data[c("DATE_START","RF")],by = "DATE_START")%>%
  arrange(CUSIP,DATE_START)
monthly_data_modified_risk_free = monthly_data_modified_risk_free[monthly_data_modified_risk_free$RET!="C",]

#calculating the excess returns over RF
monthly_data_modified_risk_free = monthly_data_modified_risk_free%>%
  mutate(across(c(RET),as.numeric),EXCESS_RETURN = RET - RF/100)


#Descriptive stats
c(count_monthly_data,mean_monthly_data,p1_monthly_data,p5_monthly_data,p25_monthly_data,p50_monthly_data, p75_monthly_data,p95_monthly_data,p99_monthly_data, std_monthly_data,skew_monthly_data,kurt_monthly_data,max_monthly_data,min_monthly_data)%<-%stats(monthly_data_modified_risk_free[c("PRC","ASKHI","BIDLO","BID","ASK","VOL","RET","SHROUT","CFACPR","CFACSHR","VWRETD","MKTCAP_ADJ","RF")])

#Descriptive stats from 1963 Onwards
c(count_monthly_data_1963,mean_monthly_data_1963,p1_monthly_data_1963,p5_monthly_data_1963,p25_monthly_data_1963,p50_monthly_data_1963, p75_monthly_data_1963,p95_monthly_data_1963,p99_monthly_data_1963, std_monthly_data_1963,skew_monthly_data_1963,kurt_monthly_data_1963,max_monthly_data_1963,min_monthly_data_1963)%<-%stats(monthly_data_modified_risk_free[monthly_data_modified_risk_free$DATE_START>="1963-01-01",c("PRC","ASKHI","BIDLO","BID","ASK","VOL","RET","SHROUT","CFACPR","CFACSHR","VWRETD","MKTCAP_ADJ","RF")])

mean_1925 = monthly_data_modified_risk_free %>%
  select(c("DATE_START","PRC","ASKHI","BIDLO","BID","ASK","VOL","RET","SHROUT","CFACPR","CFACSHR","VWRETD","MKTCAP_ADJ","RF"))%>%
  group_by(DATE_START)%>%
  mutate(RET = 100*RET,VWRETD = 100*VWRETD)%>%
  summarise(across(everything(), mean,na.rm = TRUE,.names = "mean_{.col}"))

std_1925 = monthly_data_modified_risk_free %>%
  select(c("DATE_START","PRC","ASKHI","BIDLO","BID","ASK","VOL","RET","SHROUT","CFACPR","CFACSHR","VWRETD","MKTCAP_ADJ","RF"))%>%
  group_by(DATE_START)%>%
  mutate(RET = 100*RET,VWRETD = 100*VWRETD)%>%
  summarise(across(everything(), sd,na.rm = TRUE,.names = "std_{.col}"))

#Plotting the dot plots for the mean and std monthly frequency data
plot_multiple_line_chart(mean_1925,"Mean of Variables from 1925 Onwards","$ / % for Returns")
plot_multiple_line_chart(std_1925,"Std Deviation of Variables from 1925 Onwards","$ / % for Returns")

```

##Excess Market Returns/Premium over Market Portfolio Analysis
```{r}
monthly_data_modified_risk_free = monthly_data_modified_risk_free%>%
  mutate(EXCESS_RETURN_MKT_PF = RET - VWRETD,Log_EXCESS_MKT_PF = log(1+EXCESS_RETURN_MKT_PF))


#Descriptive stats
c(count_monthly_data_mkt,mean_monthly_data_mkt,p1_monthly_data_mkt,p5_monthly_data_mkt,p25_monthly_data_mkt,p50_monthly_data_mkt, p75_monthly_data_mkt,p95_monthly_data_mkt,p99_monthly_data_mkt, std_monthly_data_mkt,skew_monthly_data_mkt,kurt_monthly_data_mkt,max_monthly_data_mkt,min_monthly_data_mkt)%<-%stats(monthly_data_modified_risk_free[c("EXCESS_RETURN_MKT_PF","Log_EXCESS_MKT_PF")])

#Descriptive stats from 1963 Onwards
c(count_monthly_data_mkt_1963,mean_monthly_data_mkt_1963,p1_monthly_data_mkt_1963,p5_monthly_data_mkt_1963,p25_monthly_data_mkt_1963,p50_monthly_data_mkt_1963, p75_monthly_data_mkt_1963,p95_monthly_data_mkt_1963,p99_monthly_data_mkt_1963, std_monthly_data_mkt_1963,skew_monthly_data_mkt_1963,kurt_monthly_data_mtk_1963,max_monthly_data_mkt_1963,min_monthly_data_mkt_1963)%<-%stats(monthly_data_modified_risk_free[monthly_data_modified_risk_free$DATE_START>="1963-01-01",c("EXCESS_RETURN_MKT_PF","Log_EXCESS_MKT_PF")])

mean_mkt_1925 = monthly_data_modified_risk_free %>%
  select(c("DATE_START","EXCESS_RETURN_MKT_PF","Log_EXCESS_MKT_PF"))%>%
  group_by(DATE_START)%>%
  mutate(EXCESS_RETURN_MKT_PF = 100*EXCESS_RETURN_MKT_PF, Log_EXCESS_MKT_PF = 100*Log_EXCESS_MKT_PF)%>%
  summarise(across(everything(), mean,na.rm = TRUE,.names = "mean_{.col}"))

std_mkt_1925 = monthly_data_modified_risk_free %>%
  select(c("DATE_START","EXCESS_RETURN_MKT_PF","Log_EXCESS_MKT_PF"))%>%
  group_by(DATE_START)%>%
  mutate(EXCESS_RETURN_MKT_PF = 100*EXCESS_RETURN_MKT_PF, Log_EXCESS_MKT_PF = 100*Log_EXCESS_MKT_PF)%>%
  summarise(across(everything(), sd,na.rm = TRUE,.names = "std_{.col}"))


plot_multiple_line_chart(mean_mkt_1925,"Mean of Excess Returns wrt Value Weighted Market Portfolio  from 1925 Onwards","%")
plot_multiple_line_chart(std_mkt_1925,"Std Deviation of Excess Returns wrt Value Weighted Market Portfolio from 1925 Onwards","%")

cumulative_excess_returns_data_monthly = monthly_data_modified_risk_free%>%
  arrange(CUSIP,DATE_START)%>%
  select(c("DATE_START","CUSIP","EXCESS_RETURN","EXCESS_RETURN_MKT_PF","Log_EXCESS_MKT_PF"))%>%
  group_by(CUSIP)%>%
  drop_na(EXCESS_RETURN,EXCESS_RETURN_MKT_PF,Log_EXCESS_MKT_PF)%>%
  mutate(COMP_EXCESS_RET = CUMULATIVE_PRODUCT(1+EXCESS_RETURN),COMP_MKT_EXCESS_RET =  CUMULATIVE_PRODUCT(1+EXCESS_RETURN_MKT_PF),CUM_MKT_LOG_EXCESS_RET = CUMULATIVE_SUM(Log_EXCESS_MKT_PF))
#Comparitive study for both time period 1925-2020 and 1963-2020
final_cumulative_data = cumulative_excess_returns_data_monthly%>%
  group_by(CUSIP)%>%
  slice(n())

final_cumulative_data_1963 = cumulative_excess_returns_data_monthly%>%
  filter(DATE_START >= "1963-01-01")%>%
  group_by(CUSIP)%>%
  slice(1,n())%>%
  mutate(COMP_EXCESS_RET = (COMP_EXCESS_RET/lag(COMP_EXCESS_RET)-1),COMP_MKT_EXCESS_RET = (COMP_MKT_EXCESS_RET/lag(COMP_MKT_EXCESS_RET)-1),CUM_MKT_LOG_EXCESS_RET = (CUM_MKT_LOG_EXCESS_RET/lag(CUM_MKT_LOG_EXCESS_RET)-1))%>%
  drop_na()


#Descriptive stats
c(count_cumulative,mean_cumulative,p1_cumulative,p5_cumulative,p25_cumulative,p50_cumulative, p75_cumulative,p95_cumulative,p99_cumulative, std_cumulative,skew_cumulative,kurt_cumulative,max_cumulative,min_cumulative)%<-%stats(final_cumulative_data[c("COMP_EXCESS_RET","COMP_MKT_EXCESS_RET","CUM_MKT_LOG_EXCESS_RET")])

#Descriptive stats from 1963 Onwards
c(count_cumulative_1963,mean_cumulative_1963,p1_cumulative_1963,p5_cumulative_1963,p25_cumulative_1963,p50_cumulative_1963, p75_cumulative_1963,p95_cumulative_1963,p99_cumulative_1963, std_cumulative_1963,skew_cumulative_1963,kurt_cumulative_1963,max_cumulative_1963,min_cumulative_1963)%<-%stats(final_cumulative_data_1963[c("COMP_EXCESS_RET","COMP_MKT_EXCESS_RET","CUM_MKT_LOG_EXCESS_RET")])


mean_cumulative_1925 = cumulative_excess_returns_data_monthly %>%
  ungroup() %>%
  select(c("DATE_START","COMP_EXCESS_RET","COMP_MKT_EXCESS_RET","CUM_MKT_LOG_EXCESS_RET"))%>%
  group_by(DATE_START)%>%
  mutate(COMP_EXCESS_RET = 100*COMP_EXCESS_RET,COMP_MKT_EXCESS_RET = 100*COMP_MKT_EXCESS_RET,CUM_MKT_LOG_EXCESS_RET = 100*CUM_MKT_LOG_EXCESS_RET)%>%
  summarise(across(everything(), mean,na.rm = TRUE,.names = "mean_{.col}"))

std_cumulative_1925 = cumulative_excess_returns_data_monthly %>%
  ungroup() %>%
  select(c("DATE_START","COMP_EXCESS_RET","COMP_MKT_EXCESS_RET","CUM_MKT_LOG_EXCESS_RET"))%>%
  group_by(DATE_START)%>%
  mutate(COMP_EXCESS_RET = 100*COMP_EXCESS_RET,COMP_MKT_EXCESS_RET = 100*COMP_MKT_EXCESS_RET,CUM_MKT_LOG_EXCESS_RET = 100*CUM_MKT_LOG_EXCESS_RET)%>%
  summarise(across(everything(), sd,na.rm = TRUE,.names = "std_{.col}"))


plot_multiple_line_chart(mean_cumulative_1925,"Mean of Cumulative Excess Returns from 1925 Onwards","%")
plot_multiple_line_chart(std_cumulative_1925,"Std Deviation of Cumulative Excess Returns from 1925 Onwards","%")
```



###Reading the Daily Data (DSF NEW)
```{r}
##Removing old variables to free up memory
rm(list = setdiff(ls(), lsf.str()))
path = "dsf_new.csv"
daily_data <- fread(path, select=c('DATE','CUSIP','PRC', 'VOL', 'RET','BID', 'ASK','ASKHI','BIDLO','OPENPRC','SHROUT','VWRETD'))

```

## Daily Data Modification, Plots for Return, Spread
```{r}
#Data Cleaning done 
daily_data <- daily_data %>%
            tibble()%>%
            mutate(RET = as.numeric(RET),PRC = abs(PRC),DATE =   as.Date(parse_date_time(DATE, orders = "Ymd")))%>%
  rename(DATE_START = DATE)

# New Variables creation
daily_data <- daily_data %>%
                  mutate(SPREAD = ASK - BID,
                         RANGE = ASKHI-BIDLO,
                         OPEN_CLOSE_SPREAD = PRC-OPENPRC)

#Descriptive Stats Calculation for overall time
c(count_daily_data,mean_daily_data,p1_daily_data,p5_daily_data,p25_daily_data,p50_daily_data, p75_daily_data,p95_daily_data,p99_daily_data, std_daily_data,skew_daily_data,kurt_daily_data,max_daily_data,min_daily_data)%<-%stats(daily_data[c("SPREAD","RANGE","OPEN_CLOSE_SPREAD","RET")])

#Mean calculation of the variables monthly frequency
mean_daily_data = daily_data %>%
  group_by(DATE_START)%>%
  mutate(RET = 100*RET)%>%
  summarise(across(c("SPREAD","RANGE","OPEN_CLOSE_SPREAD","RET"), mean,na.rm = TRUE,.names = "mean_{.col}"))
#Plotting the data
plot_multiple_line_chart(mean_daily_data,"Spread and Return","$ / % for Daily Returns")

```

## IPOs monthly plots Market Size and Retuns
```{r}
ipo_data <- daily_data %>% select(c('DATE_START', 'CUSIP', 'RET', 'VWRETD', 'PRC', 'SHROUT'))
rm(daily_data)

# Storing the data 
monthly_market_data <- ipo_data %>%
            select(c('DATE_START','VWRETD')) %>% 
            mutate(VWRETD = VWRETD + 1) %>%
            arrange(DATE_START) %>%
            distinct(DATE_START, .keep_all = TRUE) %>%
            mutate(DATE_START = floor_date(DATE_START, unit = "month")) %>%
            group_by(DATE_START) %>%
            mutate(VWRETD = cumprod(VWRETD)) %>% 
            slice(n())
#Creating the variables for the data (Yearly, monthly data)
monthly_market_data <- monthly_market_data %>% 
            ungroup()%>%
            mutate(CUM_RATE = CUMULATIVE_PRODUCT(VWRETD)) %>%
            mutate(YEARLY_MARKET_RET = LEAD(CUM_RATE,12)/CUM_RATE,
                  BY_YEARLY_MARKET_RET = LEAD(CUM_RATE,24)/CUM_RATE,
                  TRI_YEARLY_MARKET_RET = LEAD(CUM_RATE,36)/CUM_RATE) %>%
            rename(MONTHLY_RETURN = VWRETD)
# IPO CUSIPS Storing
ipo_date <- ipo_data %>% 
            select(c('DATE_START', 'CUSIP','RET')) %>%
            drop_na(RET) %>%
            select(-RET) %>%
            group_by(CUSIP) %>% 
            summarise(across(everything(), min, na.rm = TRUE)) %>% 
            inner_join(ipo_data, by = c("DATE_START", "CUSIP")) %>%
            select(c('DATE_START', 'CUSIP','RET', 'PRC','SHROUT')) %>%
            mutate(MARKET_CAP = PRC*SHROUT)
            
            
# Count of the IPOS
monthly_ipo <- ipo_date %>%
            filter((DATE_START!="1972-12-15") & (DATE_START!="1962-12-31"))%>%
            mutate(DATE_START = floor_date(DATE_START, unit = "month")) %>%
            group_by(DATE_START) %>%
            count() %>%
            inner_join(monthly_market_data %>% select(c('DATE_START','MONTHLY_RETURN')))

#Calculating the returns of the IPOs            
monthly_ipo = monthly_ipo%>%
  mutate(MONTHLY_RETURN = (MONTHLY_RETURN-1)*100)%>%
  rename(IPO_COUNT =  n)

#Line chart for the IPOs
plot_multiple_line_chart(monthly_ipo,"Monthly IPOs","Count",1,"bar")

monthly_ipo_size <- ipo_date %>%
                filter((DATE_START!="1972-12-15") & (DATE_START!="1962-12-31"))%>%
                select(c('DATE_START', 'MARKET_CAP')) %>% 
                mutate(DATE_START = floor_date(DATE_START, unit = "month")) %>%
                group_by(DATE_START) %>%
                summarise(across(everything(), mean, na.rm = TRUE)) %>% 
                inner_join(monthly_market_data %>% select(c('DATE_START','MONTHLY_RETURN')))

ipo_sz_monthly = monthly_ipo_size%>%
  mutate(MONTHLY_RETURN = (MONTHLY_RETURN-1)*100)
#Plot the Mean IPO size against the return 
plot_multiple_line_chart(ipo_sz_monthly,"Mean IPO Sizes per month","Size ($)",1)
#Calculating first day return for the IPO
monthly_first_day_return <- ipo_date %>%
                select(c('DATE_START','RET')) %>%
                 mutate(DATE_START = floor_date(DATE_START, unit = "month")) %>%
                group_by(DATE_START) %>%
                summarise(across(everything(), mean, na.rm = TRUE)) %>%
                inner_join(monthly_market_data %>% select(c('DATE_START','MONTHLY_RETURN')))
```

##IPO Data Analysis Excess Returns Market Portfolio
```{r}

#Taking all data for the IPO
all_ipo_data <- ipo_date %>% 
            select(c('DATE_START','CUSIP', 'PRC','RET')) %>%
            rename(IPO_DATE = DATE_START, IPO_PRC = PRC) %>%
            inner_join(ipo_data %>% select('DATE_START','CUSIP', 'PRC','VWRETD')) %>%
            mutate(DIFF = difftime(DATE_START,IPO_DATE,units="weeks")) %>%
            filter(DIFF == 0 | DIFF == 4 | DIFF == 52 | DIFF == 104 | DIFF == 156) %>%
            mutate(IPO_RET = ifelse(DIFF == 0, RET, (PRC-IPO_PRC)/IPO_PRC ) ) %>%
            select(c('CUSIP','DIFF','IPO_RET')) %>%
            pivot_wider(names_from = DIFF, values_from = IPO_RET) %>%
            rename( FIRST_DAY_RET = '0',
                    FIRST_MONTH_RET = '4',
                    FIRST_YEAR_MONTH = '52',
                    TWO_YEAR_RETURN = '104',
                    THREE_YEAR_RETURN = '156') %>%
            inner_join(ipo_date) %>%
            mutate(DATE_START = floor_date(DATE_START, unit = "month")) %>%
            group_by(DATE_START) %>%
            summarise(across(.cols = is.numeric, mean, na.rm = TRUE))

#Removing the IPO data(excess variables)so to free up memory space
rm(ipo_data)
#Calculating the ipo excess returns
ipo_excess_returns <- all_ipo_data %>% 
                       inner_join(monthly_market_data) %>%
                       mutate(FIRST_MONTH_EXCESS = FIRST_MONTH_RET - MONTHLY_RETURN,
                              FIRST_YEAR_EXCESS = FIRST_YEAR_MONTH - YEARLY_MARKET_RET,
                              TWO_YEAR_EXCESS = TWO_YEAR_RETURN - BY_YEARLY_MARKET_RET,
                              THREE_YEAR_EXCESS = THREE_YEAR_RETURN - TRI_YEARLY_MARKET_RET) %>%
                       select(DATE_START, FIRST_DAY_RET, FIRST_MONTH_EXCESS, FIRST_YEAR_EXCESS, TWO_YEAR_EXCESS, THREE_YEAR_EXCESS)
 
#Descriptive stats for the excess return for the IPO
c(count_ipo_excess_returns,mean_ipo_excess_returns,p1_ipo_excess_returns,p5_ipo_excess_returns,p25_ipo_excess_returns,p50_ipo_excess_returns, p75_ipo_excess_returns,p95_ipo_excess_returns,p99_ipo_excess_return, std_ipo_excess_return,skew_ipo_excess_return,kurt_ipo_excess_return,max_ipo_excess_return,min_ipo_excess_return)%<-%stats(ipo_excess_returns[,names(ipo_excess_returns)!="DATE_START"])

#Multiply by 100 for the IPO Excess Returns Columns
ipo_excess_returns[,2:ncol(ipo_excess_returns)] = ipo_excess_returns[,2:ncol(ipo_excess_returns)]*100
#Calculating mean and std for the variables(excess returns) over time (monthly frequency)
mean_ipo_excess_returns_1963 = ipo_excess_returns %>%
  group_by(DATE_START)%>%
  summarise(across(everything(), mean,na.rm = TRUE,.names = "mean_{.col}"))

std_ipo_excess_returns_1963 = ipo_excess_returns %>%
  group_by(DATE_START)%>%
  summarise(across(everything(), sd,na.rm = TRUE,.names = "std_{.col}"))

```